---
title: "BIS 15L Practice Midterm 2"
author: "Jolane Abrams"
output:
  html_document: 
    theme: spacelab
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the libraries
```{r}
library(tidyverse)
library(janitor)
library(ggthemes)
options(scipen=999) #disables scientific notation when printing
```

```{r}
library(here)
```

## Gapminder
For this assignment, we are going to use data from  [gapminder](https://www.gapminder.org/). Gapminder includes information about economics, population, social issues, and life expectancy from countries all over the world. We will use three data sets, so please load all three.  

1. population_total.csv  
2. income_per_person_gdppercapita_ppp_inflation_adjusted.csv  
3. life_expectancy_years.csv  

```{r}
population <- read_csv(here("practice_midterm2", "data", "population_total.csv")) %>% 
  mutate(country=as_factor(country))
```

```{r}
glimpse (population)
```
```{r}
income <- read_csv(here("practice_midterm2", "data", "income_per_person_gdppercapita_ppp_inflation_adjusted.csv")) %>% 
  mutate(country=as_factor(country))
```

```{r}
glimpse(income)
```

```{r}
life_expectancy <- read_csv(here("practice_midterm2", "data", "life_expectancy_years.csv")) %>% 
  mutate(country=as_factor(country))
```
```{r}
glimpse(life_expectancy)
```

1. (3 points) Once you have an idea of the structure of the data, please make each data set tidy (hint: think back to pivots) and store them as new objects. You will need both the original (wide) and tidy (long) data!  
```{r}
pop_long <- population %>%
  pivot_longer(cols = -country, names_to = "year", values_to = "population")
pop_long
```

```{r}
inc_long <- income %>%
  pivot_longer(cols = -country, names_to = "year", values_to = "income")
inc_long
```

```{r}
life_long <- life_expectancy %>%
  pivot_longer(cols = -country, names_to = "year", values_to = "life_expectancy")
life_long
```

```{r}

```

2. (1 point) How many different countries are represented in the data? Provide the total number and their names. Since each data set includes different numbers of countries, you will need to do this for each one.  
```{r}
inc_long %>% summarize(n_countries=n_distinct(country))
```

```{r}
inc_long %>% count(country) %>% select(country)
```

```{r}
life_long%>% summarize(n_countries=n_distinct(country))
```

```{r}
life_long %>% count(country) %>% select(country)
```
```{r}
pop_long%>% summarize(n_countries=n_distinct(country))
```
```{r}
pop_long %>% count(country) %>% select(country)
```

## Life Expectancy  

3. (2 points) Let's limit the data to 100 years (1920-2020). For these years, which country has the highest average life expectancy? How about the lowest average life expectancy?  

Highest
```{r}
life_long %>% 
  filter(between(year, 1920, 2020)) %>% 
  group_by(country) %>% 
  summarize(mean_life_expectancy=mean(life_expectancy)) %>% 
  arrange(desc(mean_life_expectancy))
```

Lowest
```{r}
life_long %>% 
  filter(between(year, 1920, 2020)) %>% 
  group_by(country) %>% 
  summarize(mean_life_expectancy=mean(life_expectancy)) %>% 
  arrange(mean_life_expectancy)
```

4. (3 points) Although we can see which country has the highest life expectancy for the past 100 years, we don't know which countries have changed the most. What are the top 5 countries that have experienced the biggest improvement in life expectancy between 1920-2020?  
```{r}
life_long %>% 
  select(country, year, life_expectancy) %>% 
  filter(between(year, 1920, 2020)) %>% 
  group_by(country) %>% 
  mutate(improved = last(life_expectancy) - first(life_expectancy)) %>% 
  ungroup() %>% 
  select(country, improved) %>% 
  arrange(-improved) %>% 
  distinct(country, .keep_all = TRUE) %>% 
  slice_head(n=5)
```


5. (3 points) Make a plot that shows the change over the past 100 years for the country with the biggest improvement in life expectancy. Be sure to add appropriate aesthetics to make the plot clean and clear. Once you have made the plot, do a little internet searching and see if you can discover what historical event may have contributed to this remarkable change.  

```{r}
kuwait <- life_long %>% 
filter(country=="Kuwait") %>% 
  filter(between(year, 1920, 2020)) %>%
  select(year, life_expectancy)%>% 
  ggplot(aes(x=year, y=life_expectancy, group=1))+
  geom_line(color="green")
  
```

```{r}
kuwait+
  theme_gray()+
  theme(plot.title = element_text(size = 12, face = "bold"),
        axis.title = element_text(size = 8),
        axis.text.x = element_text(angle=85, hjust = 1, size=5)) +
  labs(title = "Life Expectancy in Kuwait (1920-2020)",
       x = NULL,
       y = "Life Expectancy (years)")
```

## Population Growth
6. (3 points) Which 5 countries have had the highest population growth over the past 100 years (1920-2020)?  
```{r}
pop_long %>% 
  filter(between(year, 1920, 2020)) %>% 
  filter(year=="1920" | year=="2020") %>% 
  pivot_wider(names_from = year,
              values_from = population) %>% 
  mutate(growth =`2020`-`1920`) %>% 
  top_n(5, growth) %>% 
  arrange(-growth)
```

7. (4 points) Produce a plot that shows the 5 countries that have had the highest population growth over the past 100 years (1920-2020). Which countries appear to have had exponential growth?  

```{r}
population %>% 
  filter(country=="India" | country=="China" | country=="Pakistan" | country=="Indonesia" | country=="United States") %>% 
  pivot_longer(-country,
               names_to = "year",
               values_to="population") %>% 
  filter(year>=1920 & year<=2020) %>% 
  ggplot(aes(x=year, y=population, group=country, color=country))+
  geom_line()+
  theme_linedraw()+
  theme(plot.title = element_text(size = 12, face = "bold"),
        axis.title = element_text(size = 10),
        axis.text.x = element_text(angle=75, hjust = 1, size=4)) +
  labs(title = "Population Growth in the 5 Highest-Growth Countries, 1920-2020",
       x = NULL,
       y = "Population (millions)")

```

```{r}
```


## Income
The units used for income are gross domestic product per person adjusted for differences in purchasing power in international dollars.

8. (4 points) As in the previous questions, which countries have experienced the biggest growth in per person GDP. Show this as a table and then plot the changes for the top 5 countries. With a bit of research, you should be able to explain the dramatic downturns of the wealthiest economies that occurred during the 1980's.  
```{r}
inc_long %>% 
  filter(between(year, 1920, 2020)) %>% 
  filter(year=="1920" | year=="2020") %>% 
  pivot_wider(names_from = year,
              values_from = income) %>% 
  mutate(growth =`2020`-`1920`) %>% 
  top_n(5, growth) %>% 
  arrange(-growth)
```

```{r}
income %>% 
  filter(country=="Qatar" | country=="Luxembourg" | country=="Singapore" | country=="Brunei" | country=="Ireland") %>% 
  pivot_longer(-country,
               names_to = "year",
               values_to="income") %>% 
  filter(year>=1920 & year<=2020) %>% 
  ggplot(aes(x=year, y=income, group=country, color=country))+
  geom_line()+
  theme_linedraw()+
  theme(plot.title = element_text(size = 12, face = "bold"),
        axis.title = element_text(size = 10),
        axis.text.x = element_text(angle=75, hjust = 1, size=4)) +
  labs(title = "Per Capita GDP Growth in the 5 Highest-Growth Countries, 1920-2020",
       x = NULL,
       y = "Per Capita GDP (international dollars)")
```
OPEC monopolized oil production in the 1970s, raising oil prices sky-high, which led to the massive GDP increases in Qatar and Brunei.


9. (3 points) Create three new objects that restrict each data set (life expectancy, population, income) to the years 1920-2020. Hint: I suggest doing this with the long form of your data. Once this is done, merge all three data sets using the code I provide below. You may need to adjust the code depending on how you have named your objects. I called mine `life_expectancy_100`, `population_100`, and `income_100`. For some of you, learning these `joins` will be important for your project.  
```{r}
pop100 <- pop_long %>% 
   filter(between(year, 1920, 2020))
pop100
```

```{r}
inc100 <- inc_long %>% 
  filter(between(year, 1920, 2020))
inc100
```

```{r}
life100 <- life_long %>% 
  filter(between(year, 1920, 2020))
life100
```

```{r}
gapminder_join <- inner_join(life100, pop100, by= c("country", "year"))
gapminder_join <- inner_join(gapminder_join, inc100, by= c("country", "year"))
gapminder_join
```

10. (4 points) Use the joined data to perform an analysis of your choice. The analysis should include a comparison between two or more of the variables `life_expectancy`, `population`, or `income.`  

```{r}
gapminder_join %>% 
  group_by(country) %>% 
  summarize(mean_life_expectancy=mean(life_expectancy, na.rm=T),
            mean_income=mean(income, na.rm=T),
            mean_population=mean(population, na.rm=T))
```

```{r}
gapminder_join %>% 
  filter(between(year, 1970, 2020)) %>% 
  group_by(country) %>% 
  summarize(mean_life_expectancy=mean(life_expectancy, na.rm=T),
            mean_income=mean(income, na.rm=T)) %>% 
  ggplot(aes(x=mean_life_expectancy, y=log10(mean_income)))+
  geom_point()+
  geom_smooth(method=lm, se=F)+
  theme_linedraw()+
  theme(plot.title = element_text(size = 12, face = "bold"),
        axis.title = element_text(size = 10),
        axis.text.x = element_text(size=10)) +
  labs(title = "Log10 GDP vs. Life Expectancy, 1970 - 2020",
       x = "Life Expectancy (years)",
       y = "Log10 GDP")
```