---
title: "Lab 4 Homework"
author: "Jolane Abrams"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the tidyverse
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

## Data
For the homework, we will use data about vertebrate home range sizes. The data are in the class folder, but the reference is below.  

**Database of vertebrate home range sizes.**  
Reference: Tamburello N, Cote IM, Dulvy NK (2015) Energy and the scaling of animal space use. The American Naturalist 186(2):196-211. http://dx.doi.org/10.1086/682070.  
Data: http://datadryad.org/resource/doi:10.5061/dryad.q5j65/1  

**1. Load the data into a new object called `homerange`.**
```{r}
homerange <- readr::read_csv("data/Tamburelloetal_HomeRangeDatabase.csv")
```


**2. Explore the data. Show the dimensions, column names, classes for each variable, and a statistical summary. Keep these as separate code chunks.** 
```{r}
dim (homerange)
```

```{r}
glimpse(homerange)
```
```{r}
str(homerange)
```


```{r}
view(homerange)
```


```{r}
colnames(homerange)
```
```{r}
summary(homerange)
```


**3. Change the class of the variables `taxon` and `order` to factors and display their levels.**  
```{r}
homerange$taxon <- as.factor(homerange$taxon)
class (homerange$taxon)
homerange$order <- as.factor(homerange$order)
class (homerange$order)
```
```{r}
levels(homerange$taxon)
```
```{r}
levels (homerange$order)
```



**4. What taxa are represented in the `homerange` data frame? Make a new data frame `taxa` that is restricted to taxon, common name, class, order, family, genus, species.**  

```{r}
taxa <- select(homerange, "taxon", "common.name", "class","order", "family", "genus", "species")
taxa
```

**5. The variable `taxon` identifies the large, common name groups of the species represented in `homerange`. Make a table the shows the counts for each of these `taxon`.** 

```{r}
table(homerange$taxon)
```



**6. The species in `homerange` are also classified into trophic guilds. How many species are represented in each trophic guild.**  

```{r}
table(homerange$trophic.guild)
```


**7. Make two new data frames, one which is restricted to carnivores and another that is restricted to herbivores.**  
```{r}
carn <- filter(homerange, trophic.guild=="carnivore")
carn
```
```{r}
herb <- filter(homerange, trophic.guild=="herbivore")
herb
```

**8. Do herbivores or carnivores have, on average, a larger `mean.hra.m2`? Remove any NAs from the data.**  
```{r}
anyNA(homerange)
```
```{r}
mean(herb$mean.hra.m2, na.rm=T)
```
```{r}
mean(carn$mean.hra.m2, na.rm=T)
```
Herbivores have a larger average mean.hra.m2.


**9. Make a new dataframe `deer` that is limited to the mean mass, log10 mass, family, genus, and species of deer in the database. The family for deer is cervidae. Arrange the data in descending order by log10 mass. Which is the largest deer? What is its common name?**  

```{r}
deer <-select(homerange,"mean.mass.g","log10.mass","family","genus","species")
deer
```
```{r}
deer <- filter(deer, family=="cervidae")
deer
```

```{r}
arrange (deer, -log10.mass)
```
```{r}
filter(homerange, species=="alces")
```
Moose (Alces alces) is the largest deer species in the data set.

**10. As measured by the data, which snake species has the smallest homerange? Show all of your work, please. Look this species up online and tell me about it!** **Snake is found in taxon column**    

```{r}
snek <- filter(homerange, taxon =="snakes")
snek
```
```{r}
min (snek$mean.hra.m2)
```
```{r}
filter(snek, mean.hra.m2==200)
```
According to Wikipedia (https://en.wikipedia.org/wiki/Bitis_schneideri), Bitis schneiderei (aka the namaqua dwarf adder) is a venomous snake native to a small coastal region that straddles the border between Namibia and South Africa. B. schneideri is possibly the world's smallest viper, with an average total length (including tail) of 18–25 cm (7–10 in) and a maximum reported total length of 28 cm (11 in). Its venom causes symptoms including intense pain, pronounced swelling, discoloration and oozing of serum from the punctures. After 24 hours, a hematoma may develop at the bite site. Victim's condition stabilized after 3-4 days, with the swelling and pain gradually subsiding. Healing was complete after two weeks, with no loss of function or sensation of the afflicted digit. Minimal local tissue damage and no systemic effects were reported. No antivenin is available for this species.




## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.   