Based on 2022 paper, "https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0278328" Does culture moderate relationships between rumination and symptoms of depression and PTSD? 
Brooding/RRS rates & culture with: 

Akshanth:      religion, gender, education(1-5; change to factors) 

JKA:           depression, anxiety, age, trauma type for PTSD


----------------------------------------------------------------------------------------------

Data codes for variables of interest from paper:

For both: First 109 participant ids are Euro-Australian, next 144 are Malaysian

For depression data:

2 Cultures:  0,1 = Euro-Australian, Malaysian (convert to factors)

Gender (convert to factors):              
 M = 1
 F = 2
Declined to state = 5  

Education (Highest level achieved):
Secondary = 1
Post-secondary = 2
Undergrad degree = 3
Graduate degree = 4
Other = 5

Religion codes:
None = 1
Christian = 2
Muslim = 3
Buddhist/Taoist = 4
Hindu = 5
Other (incl Sikh, Agnostic) = 6

PTSD dataset:

Culture = Groups 1,2  = Euro-Australian, Malaysian (change to factors)

Index trauma type (what the PTSD is thought to be in response to):

Accident/serious injury/illness = 1
Non-sexual assault/abuse = 2
Sexual assault/abuse = 3
Witnessing death = 4
War/natural disaster = 5
Other = 6       

HADS = Hospital Anxiety and Depression Scale: 

had_sdep = depression subscale total
had_sanx = anxiety subscale total
had_stotal = overall HADS total

pts_dtotal = overall score on PTSD assessment scale

RRS-B = Brooding scale - want sum of all RRS scores

```{r}
library(here)
library(tidyverse)
library(naniar)
library(janitor)
library(skimr)
library(ggplot2)
library(ggthemes)
library(RColorBrewer)
library(paletteer)
library(tidyr)
```

```{r}
my_palette <- paletteer::paletteer_d(("ggprism::colors"))
```

```{r}
dep<-read_csv(here("jka_r_project", "rum_depress_jka.csv")) %>% clean_names
```

```{r}
ptsd<-read_csv(here("jka_r_project", "rum_ptsd_jka.csv")) %>% clean_names
```

```{r}
glimpse (dep)
```

````{r}
dep$culture <- as.factor(dep$culture)
class(dep$culture)
```


```{r}
dep %>% 
  filter(culture==0) %>% 
  summarize(max_age = max(age, na.rm =T),
            min_age = min(age,na.rm =T),
            avg_age = mean(age,na.rm =T),
            total = n())
```

```{r}
dep %>% 
  filter(culture==1) %>% 
  summarize(max_age = max(age, na.rm =T),
            min_age = min(age,na.rm =T),
            avg_age = mean(age,na.rm =T),
            total = n())
```
```{r}
dep <- dep %>% 
  mutate(age_cat = case_when(age < 21 ~ "18 - 21",
                             age >= 21 & age < 25 ~ "21 - 25",
                             age >= 25 & age < 30 ~ "25 - 30",
                             age >= 30 & age < 35 ~ "30 - 35",
                             age >= 35 & age < 40 ~ "35 - 40",
                             age >= 40 & age < 45 ~ "40 - 45",
                             age >= 45 & age < 50 ~ "45 - 50",
                             age >= 50 & age < 55 ~ "50 - 55",
                             age >= 55 & age < 60 ~ "55 - 60",
                             age >= 60 ~ "Over 60"))

```


```{r}
class(dep$age_cat)
```


```{r}
dpage <- dep %>%
#  na.omit() %>%
  group_by(culture, age_cat) %>%
  summarize(avg_brood = mean(brooding_total, na.rm = T), .groups ="keep") %>%
  select(culture, age_cat, avg_brood)
  
dpage$age_cat <- factor(dpage$age_cat, levels=c(
    "18 - 21",
    "21 - 25",
    "25 - 30",
    "30 - 35",
    "35 - 40",
    "40 - 45",
    "45 - 50",
    "50 - 55",
    "55 - 60",
    "Over 60"))

dpage
```

reorder the column order by age, 
and make them the same scale (0-14)


```{r}
dpage %>% 
  filter(!is.na(age_cat)) %>% 
  ggplot(aes(x = age_cat, y = avg_brood, fill = age_cat)) +
  geom_col(alpha = 0.5, color = "black")+
  facet_grid(culture ~ .,
             scales = "free_y", 
             space = "free", 
             labeller = labeller(culture = c(`0` = "Euro-Australian", `1` = "Malaysian"))) +
  scale_x_discrete(limits=c(
    "18 - 21",
    "21 - 25",
    "25 - 30",
    "30 - 35",
    "35 - 40",
    "40 - 45",
    "45 - 50",
    "50 - 55",
    "55 - 60",
    "Over 60"))+
  ylim(0,14)+
  labs(title = "Brooding by Age Category and Culture",
      x = NULL,
      y = "Average Brooding Score",
      fill = "Age Category") +
  theme(plot.title = element_text(size = 12, face = "bold"),
  plot.subtitle = element_text(size = 10),
  axis.title = element_text(size = 10),
  axis.text.x = element_text(angle = 45, hjust = 1))
```



```{r}
dp2 <- dep %>%
group_by(culture, depression_total) %>%
  summarize(avg_brood = mean(brooding_total, na.rm = T), .groups ="keep") %>%
  select(culture, depression_total, avg_brood)
  
dp2$depression_total <- factor(dp2$depression_total)
```

```{r}
dp2 %>% 
  filter(!is.na(depression_total)) %>% 
  ggplot(aes(x = depression_total, y = avg_brood, fill = depression_total)) +
  geom_col(alpha = 0.5, color = "black")+
  facet_grid(culture ~ .,
             scales = "free_y", 
             space = "free", 
             labeller = labeller(culture = c(`0` = "Euro-Australian", `1` = "Malaysian"))) 
```


```{r}
dp2 %>% 
  filter(!is.na(depression_total)) %>% 
  ggplot(aes(x = depression_total, y = avg_brood, fill = depression_total)) +
  geom_col(alpha = 0.5, color = "black")+
  facet_grid(culture ~ .,
             scales = "free_y", 
             space = "free", 
             labeller = labeller(culture = c(`0` = "Euro-Australian", `1` = "Malaysian")))+ 
  ylim(0,18)+
labs(title = "Brooding by Depression Level and Culture",
       x = "Total Depression Score",
       y = "Average Brooding Score",
       fill = "Total Depression Score") + 
  theme(plot.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 10),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 8))
```


```{r}
danx <- dep %>% 
  group_by(culture, anxiety_total) %>% 
  summarize(avg_brood = mean(brooding_total, na.rm = T), .groups ="keep")
  

danx$anxiety_total <- factor(danx$anxiety_total)


```

```{r}
danx <- danx %>% 
  filter(!is.na(anxiety_total) & 
         !is.na(avg_brood) & 
         !is.na(culture))

danx <- danx[-35,]

ggplot(danx, aes(x = anxiety_total, y = avg_brood, fill = anxiety_total)) +
  geom_col(alpha = 0.5, color = "black") +
  facet_grid(culture ~ .,
             scales = "free_y", 
             space = "free", 
             labeller = labeller(culture = c(`0` = "Euro-Australian", `1` = "Malaysian")))+ 
  scale_y_continuous(limits = c(0, 20))+
   labs(title = "Brooding by Anxiety Level and Culture",
  x = "Total Anxiety Score",
  y = "Average Brooding Score",
  fill = "Total Anxiety Score") + 
  theme(plot.title = element_text(size = 12, face = "bold"),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8))
```


```{r}
```

```{r}
```

```{r}
glimpse(ptsd)
```
```{r}
ptsd_t1 <- ptsd %>% 
  filter(between(as.numeric(as.character(traumacoding)), 1, 6))


  ptsd_t1 %>% 
    select(traumacoding)

```

```{r}
ptsd_t1 %>% 
  mutate(index_t = case_when (traumacoding == 1 ~ "AI",
                             traumacoding == 2 ~ "NSA",
                             traumacoding == 3 ~ "SA",
                             traumacoding == 4 ~ "CD",
                             traumacoding == 5 ~ "WND",
                             traumacoding == 6 ~ "OTH")) %>% 
  select(traumacoding, index_t)
```



```{r}
ptsd$rrs <- factor(ptsd$rrs)
ptsd$group <- factor(ptsd$group)
ptsd$had_sdep <- factor(ptsd$had_sdep)
ptsd$had_sanx <- factor(ptsd$had_sanx)
ptsd$had_stotal <- factor(ptsd$had_stotal)
ptsd$pts_dtotal  <- factor(ptsd$pts_dtotal)
ptsd$traumacoding  <- factor(ptsd$traumacoding)
class(rrs, group)

```

```{r}
names(ptsd)
```
```{r}
ptsd %>%
group_by(group, had_sdep, rrs) %>%
  summarize(avg_brood = mean(rrs, na.rm = T), .groups ="keep") %>%
  #select(group, had_sdep, rrs)
  ggplot(aes(x = had_sdep, y = rrs, fill = had_sdep)) +
  geom_col(alpha = 0.5, color = "black")
#facet_grid(group ~ .,
             #scales = "free_y", 
             #space = "free", 
             #labeller = labeller(culture = c(`0` = "Euro-Australian", `1` = "Malaysian"))) 
```

