---
title: "Lab 12 Homework: Wolves"
author: "Jolane Abrams"
date: "2023-03-02"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(janitor)
library(shiny)
library(shinydashboard)
library(ggthemes)
```


## Practice
Let's try to build an app that allows users to explore sexual composition of wolf packs by `pop`. These data are from: Brandell, Ellen E (2021), Serological dataset and R code for: Patterns and processes of pathogen exposure in gray wolves across North America, Dryad, [Dataset](https://doi.org/10.5061/dryad.5hqbzkh51)  

1. Start by loading and exploring the data.
```{r}
wolves <- read_csv("data/wolves_data/wolves_dataset.csv")
```

```{r}
names(wolves)
```


```{r}
glimpse(wolves)
```

2. For the app, here is a faceted version of what we are looking for. We want the reactive part to be `pop` and you should use `shinydashboard`.
```{r}
wolves %>% 
  filter(sex!="NA") %>% 
  ggplot(aes(x=sex, fill=sex))+
  geom_bar()+
  facet_wrap(~pop)
```


```{r}
library(shiny)
ui <- fluidPage(titlePanel("Sexual Composition of Wolf Packs by Population"),
                selectInput("x", "Select Population", choices=c(unique(wolves$pop)),
                selected="AK.PEN"),
                plotOutput("plot", width="600px", height="500px")
)
server <- function(input, output, session) {
  output$plot <- renderPlot({
    wolves %>% 
     filter(pop==input$x) %>% 
      filter(sex!="NA") %>% 
    ggplot(aes(x=sex, fill=sex)) +
      geom_bar(alpha=0.5, color="black") +
      theme_light(base_size=18) +
      labs(x=NULL, y=NULL, fill="Sex")
  })
   session$onSessionEnded(stopApp)
}
shinyApp(ui = ui, server = server)
```

```{r}
library(shinydashboard)
 ui <- dashboardPage(
   dashboardHeader(title = "Sexual Composition of Wolf Packs by Population"),
   dashboardSidebar(disable=T),
   dashboardBody(
                selectInput("x", "Select Population", choices=c(unique(wolves$pop)),
                 selected="AK.PEN"),
                 plotOutput("plot", width="600px", height="500px"))
 )
 server <- function(input, output, session) {
   output$plot <- renderPlot({
     wolves %>% 
      filter(sex!="NA") %>% 
       filter(pop==input$x) %>% 
     ggplot(aes(x=sex, fill=sex)) +
       geom_bar(alpha=0.5, color="black") +
       theme_light(base_size=18) +
       labs(x=NULL, y=NULL, fill="Sex")
   })
  session$onSessionEnded(stopApp)
 }
shinyApp(ui = ui, server = server)